---
title: "Parameter Estimation – Wild Fish Catch"
author: "Catherine Takata"
date: "3/2/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: show
    toc: true
---

## Summary 
We will identify an equation with parameters estimated by nonlinear least squares for the increase in global wild fish catch from 1950 – 2012.

**Source**: Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(boot)
library(patchwork)
library(broom)
library(nlstools)
```

```{r}
# Read in data 
fish_catch <- read.csv(here("a4_task2", "data", "fish_catch.csv"), col.names = c("year", "wild_catch", "farmed_fish", "total_fish_production")) %>% 
  mutate(time = year) 

fish_catch[1, 5] = 0
```


## Exploratory graph 

```{r}
ggplot(data = fish_catch, aes(x = year, y = wild_catch)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Wild catch from 1950-2012",
       x = "time (years)", 
       y = "wild catch (million tons)") 
```

```{r}
# Look at the log transformed data:
ggplot(data = fish_catch, aes(x = year, y = log(wild_catch))) +
  geom_point() +
  theme_minimal() +
  labs(title = "Logrithmic wild catch from 1950-2012",
       x = "time (years)", 
       y = "ln(wild catch)") 
```

## Parameter outcomes 






## Finalized results 