---
title: "Parameter Estimation of Wild Fish Catch"
author: "Catherine Takata"
date: "3/2/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: show
    toc: true
---

## Summary 
We will identify an equation with parameters estimated by nonlinear least squares for the increase in global wild fish catch from 1950 – 2012.

**Source**: Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(boot)
library(broom)
library(kableExtra)
```

```{r}
# Read in data 
fish_catch <- read.csv(here("a4_task2", "data", "fish_catch.csv"), col.names = c("year", "wild_catch", "farmed_fish", "total_fish_production")) %>% 
  mutate(time = year) 

fish_catch[1:63, 5] = c(0:62)
```


## Exploratory Data Visualization 

```{r}
ggplot(data = fish_catch, aes(x = year, y = wild_catch)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Wild catch from 1950-2012",
       x = "time (years)", 
       y = "wild catch (million tons)") 
```

```{r, include = F, message = F}
ggplot(data = fish_catch, aes(x = year, y = log(wild_catch))) +
  geom_point() +
  theme_minimal() +
  labs(title = "Wild catch from 1950-2012",
       x = "time (years)", 
       y = "ln(wild catch)") 
```


Our initial exploration reveals a trend simialr to logistic growth of a population over time. As the number of wild catch increase with time, resource limitations halt growth, revealing a carry capacity shown at about 95 million tons of wild catch. Mathematically we can identiify paramters of K and A of the growth equation for the population at time t where: 

$$A = \frac{K- P0}{P0}$$
By identifying the relationship between K and A we can further evalute the growth curve for: 


$$P(t) = \frac{K}{1+Ae^{-kt}}$$



Our original start estimates of K come from the point at which resrouces stop supporting growth of population at the carring capacity where graph levels out at a constant rate. We can then calculate A as we relate the intial population of ~17 million tons to the carrying capacity. By using non linear least sqaures, we will find parameters to minimize square residuals by using iterative algorithms to converge on parameter estimates (since there is no single equation to calculate exactly what the paramter estimates would be). 


## Inital Estimates 

```{r, include=F, message=F}
# Initial Estimates 

# Estimate the growth rate constant (k) *during exponential growth phase)
fish_est <- fish_catch %>%
  filter(wild_catch < 95) %>% 
  mutate(ln_wild_catch = log(wild_catch))

# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant
lm_k <- lm(ln_wild_catch ~ time, data = fish_est)
lm_k
```

We have initial estimate for k (0.03), and we can estimate K ~ 95 and A ~ 4.3.


## Finalized Parameter Results 

```{r, include=F, message=F}
fish_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*time)),
              data = fish_catch,
              start = list(K = 95, A = 4.3, r = 0.03),
              trace = TRUE
              )
```

```{r, include = F, message=F}
#See the model summary (null hypothesis: parameter value = 0)
summary(fish_nls)

# Use broom:: functions to get model outputs in tidier format: 
model_out <- broom::tidy(fish_nls) %>% 
  mutate(p.value = case_when(
    p.value < 0.0001 ~ "<0.0001"
  ))

# Gets the first rown 
A_est <- model_out$estimate[2] 
```

**Table 1.** Model parameters outcomes (growth rate constant (k), carrying capacity (K), and (A)) for recorded wild fish catch. 

```{r}
model_out %>% 
  kbl(col.names = c("Parameter", "Estimate", "Standard Error", "Statistic", "p - value")) %>% 
  # kbl/kable() creates a clean data table for the knitted document  with labeled columns 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```




$$P(t) = \frac{100.3}{1+4.3e^{-0.06t}}$$

```{r}
# Make predictions for the population at all of those times (time) in the original df: 
p_predict <- predict(fish_nls)

# Bind predictions to original data frame:
fish_complete <- data.frame(fish_catch, p_predict)
```


```{r}
# Plot them all together:
ggplot(data = fish_complete, aes(x = year, y = wild_catch)) +
  geom_point() +
  geom_line(aes(x = year, y = p_predict)) +
  theme_minimal() + 
  labs(title = "Least Squares Model of Wild Fish Catch Logistic Growth \nfrom 1950-2012",
       x = "Years",
       y = "Wild Catch (millions of tons)",
       caption = "Figure 1")
```


**Figure 1.** Non linear least square paramter calculations of wild catch fish from 1950-2012. Points correspond to original observations of logistic growth. Superimposed linear model indicates the paramters of k, K, and A to best fit our actual data based on interative algorithms convergered on parameter estimates. Data: Earth Policy Institute (1950-2010)





