---
title: "Bootstrapped Confidence Interval for Proportions"
author: "Catherine Takata"
date: "3/3/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: show
    toc: true
---

## Summary 
We will find the bootstrapped 95% confidence interval for the proportion of genderqueer individuals experiencing exclusionary conduct based on the 2014 UCSB survey data.
Bootstrapping uses **sampling with replacement** to find a sampling distribution that is based on more than a single sample.

Source: 2014 UCSB Campus Climate Project Final Report (prepared by Rankin & Associates Consulting. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(here)
library(boot)
library(patchwork)
library(broom)
library(nlstools)
```

## Builds vector data 

```{r}
identified_exlusion <- rep(c(1, 0),times = c(22, 14))

gender_df <-data.frame(identified_exlusion)
```

## Bootstrapping 

```{r}
# Create a function that will calculate the proportion of each bootstrapped sample
prop_fun <- function(x,i) {sum(x[i]==1)/length(x)}

# Create 100 bootstrap samples by resampling from the identified exclusion vector, using the function you created (prop_fun) to calculate the proportion of each:
boot_100 <- boot(identified_exlusion, 
                 statistic = prop_fun,
                 R = 100)

# Create 10000 bootstrap samples with replacement
boot_10k <- boot(identified_exlusion,
                 statistic = prop_fun, 
                 R = 10000)

```

## Histogram 

```{r}
# Make vectors of bootstrap sample proportions
# Create a data frame of the vectors to use ggplot 
boot_100_df <- data.frame(bs_prop = boot_100$t)
boot_10k_df <- data.frame(bs_prop = boot_10k$t)
```

```{r}
# Histogram of 100 bootstrap sample means:
plot_10k <- ggplot(data = boot_10k_df, aes(x = bs_prop)) +
  geom_histogram() +
  theme_minimal() +
  labs(title = "Distribution of Proportion of Nonbinary/Genderqueer Students \nthat Experienced Exclusion",
       x = "Proportion of Students", 
       y = "Counts",
       caption = "Figure 1")

plot_10k
```

**Figure 1**: Sampling distribution based on proportions calculated from 10,000 bootstrap samples. Students who identified as genderqueer individuals valued at [1] to indicate "YES" to experiencing exclusionary conduct based on the 2014 UCSB survey data. 


## Confidence Interval

```{r}
# Use `boot.ci()` to find the confidence interval for the bootstrapped distribution (here, with the 10k bootstrapped means):
boot.ci(boot_10k, conf = 0.95, type = "perc")
```

There is a 95% confidence interval that the 0.44 to 0.78 range contains the true proportion of nonbinary/genderqueer students who experienced exlusion based on 10,000 bootrap generated samples of replacement. 


