---
title: "Agglomerative Hierarchical Clustering of Santa Barbara Stream Chemistry"
author: "Catherine Takata"
date: "3/2/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: show
    toc: true
---

## Summary 
We will perform hierarchical clustering by site. First, we will make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

```{r}
# Read in data and convert -999 values to NA.
water_chem <- read.csv(here("a4_task1", "data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  na_if(-999.0) %>% 
  clean_names()
```

```{r}
# Take means from all water chem variables (observations) per site
water_chem$summary <- rowMeans(water_chem[3:12], na.rm = TRUE)

# Create a data set that has a single summary row per site 
water_sum <- water_chem %>% 
  select(site_code, summary) %>% 
  group_by(site_code) %>% 
  summarise(mean = mean(summary))
```

```{r}
# Scale the numeric variables 
water_scaled <- water_sum %>% 
  select(2) %>% 
  scale()
  
# Update to add rownames (site_code) from water_chem
rownames(water_scaled) <- water_sum$site_code

```

## Euclidean distances
Use the `stats::dist()` function to find the Euclidean distance in multivariate space between the different water quality indexes (by site):

```{r}
# Compute dissimilarity values (Euclidean distances):
euc_distance <- dist(water_scaled, method = "euclidean")
```

## Complete Linkage Dendrogram
Use stats::hclust() function to calcualte hierarchical clustering based on a matrix of euclidean distances. 
Use complete linkages - merge clusters by the smallest maximum distance between two observations in distinct clusters. 

```{r, fig.show = 'hide'}
# Hierarchical clustering (complete linkage)
cluster_complete <- hclust(euc_distance, method = "complete" )

# Plot it (base plot):
plot(cluster_complete, cex = 0.6, hang = -1)
```

```{r}
ggdendrogram(cluster_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(title = "Hierarchical Clustering of Santa Barbara Stream Chemistry by Site",
       x = "Sites",
       y = "")
```


